---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.align="center"
)
library("rjd3workspace")
```

# rjd3workspace


**rjd3workspace** offers several functions to wrangle JDemetra+ v3.x workspaces.

Seasonal adjustment with X-12ARIMA can be done with the package [**rjd3x13**](https://github.com/rjdverse/rjd3x13) and with TRAMO-SEATS with the package [**rjd3tramoseats**](https://github.com/rjdverse/rjd3tramoseats).


## Installation

**rjd3workspace** relies on the [**rJava**](https://CRAN.R-project.org/package=rJava) package 

Running rjd3 packages requires **Java 17 or higher**. How to set up such a configuration in R is explained [here](https://jdemetra-new-documentation.netlify.app/#Rconfig)

To get the current stable version (from the latest release):

```{r, eval = FALSE}
# install.packages
remotes::install_github("rjdverse/rjd3toolkit@*release")
remotes::install_github("rjdverse/rjd3tramoseats@*release")
remotes::install_github("rjdverse/rjd3x13@*release")
remotes::install_github("rjdverse/rjd3providers@*release")
remotes::install_github("rjdverse/rjd3workspace@*release")
```

To get the current development version from GitHub:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("rjdverse/rjd3workspace")
```


## Usage

```{r, echo = TRUE, eval = TRUE}
library("rjd3workspace")

dir <- tempdir()

y <- rjd3toolkit::ABS$X0.2.09.10.M
jws <- .jws_new()
jsap1 <- .jws_sap_new(jws, "sa1")
add_sa_item(jsap1, name = "x13", x = rjd3x13::x13(y))
add_sa_item(jsap1, name = "tramo", x = rjd3tramoseats::tramoseats(y))
save_workspace(jws, file.path(dir, "ws.xml"))

jws <- .jws_load(file = file.path(dir, "ws.xml"))
.jws_compute(jws) # to compute the models
jsap1 <- .jws_sap(jws, idx = 1) # first SAProcessing
jsa1 <- .jsap_sa(jsap1, idx = 1) # first SAItem
.jsa_name(jsa1)
mod1 <- .jsa_read(jsa1)
```


### Create SA-item with path

``` {r create_sa_item_path, echo = TRUE, eval = TRUE}
# install.packages("remotes")
# remotes::install_github("rjdverse/rjd3providers")

dir <- tempdir()

xlsx_file <- paste0(system.file("examples", package="rjd3providers"), "/Insee.xlsx")
ts1 <- rjd3providers::spreadsheet_series(file = xlsx_file, sheet = 1L, series = 3L)

y <- ts1$data
jws <- .jws_new()
jsap1 <- .jws_sap_new(jws, "sa1")
add_sa_item(jsap1, name = "x13", x = rjd3x13::x13(y))
rjd3workspace::set_ts(jsap = jsap1, idx = 1L, y = ts1)
add_sa_item(jsap1, name = "tramo", x = rjd3tramoseats::tramoseats(y))
rjd3workspace::set_ts(jsap = jsap1, idx = 2L, y = ts1)

save_workspace(jws, file.path(dir, "ws.xml"))

jws <- .jws_load(file = file.path(dir, "ws.xml"))
.jws_compute(jws) # to compute the models
jsap1 <- .jws_sap(jws, idx = 1) # first SAProcessing
jsa1 <- .jsap_sa(jsap1, idx = 1) # first SAItem
.jsa_name(jsa1)
mod1 <- .jsa_read(jsa1)
```


## Package Maintenance and contributing

Any contribution is welcome and should be done through pull requests and/or issues.
pull requests should include **updated tests** and **updated documentation**. If functionality is changed, docstrings should be added or updated.


## Licensing

The code of this project is licensed under the [European Union Public Licence (EUPL)](https://joinup.ec.europa.eu/collection/eupl/eupl-text-eupl-12).
