---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.align="center"
)
library(rjdemetra3)
```

# RJDemetra3


RJDemetra3 offers several functions to interact with JDemetra+ v3.0 workspaces.
Seasonal adjustment with X-12ARIMA can be done with the package [rjd3x13](https://github.com/rjdemetra/rjd3x13) and with TRAMO-SEATS with the package [rjd3tramoseats](https://github.com/rjdemetra/rjd3tramoseats).

## Installation

RJDemetra3 relies on the [rJava](https://CRAN.R-project.org/package=rJava) package and Java SE 17 or later version is required.


```{r, eval = FALSE}
# Install development version from GitHub
# install.packages("remotes")
remotes::install_github("rjdemetra/rjd3toolkit")
remotes::install_github("rjdemetra/rjd3x13")
remotes::install_github("rjdemetra/rjd3tramoseats")
remotes::install_github("rjdemetra/rjdemetra3")
```

## Usage

RJDemetra3 relies on the [rJava](https://CRAN.R-project.org/package=rJava) package and Java SE 17 or later version is required.


```{r}
library(rjdemetra3)
dir <- tempdir()
y <- rjd3toolkit::ABS$X0.2.09.10.M
jws <- .jws_new()
jsap1 <- .jws_sap_new(jws, "sa1")
add_sa_item(jsap1, name = "x13", x = rjd3x13::x13(y))
add_sa_item(jsap1, name = "tramo", x = rjd3tramoseats::tramoseats(y))
save_workspace(jws, file.path(dir, "wk.xml"))

jws <- .jws_open(file = file.path(dir, "wk.xml"))
.jws_compute(jws) # to compute the models
jsap1 <- .jws_sap(jws, idx = 1) # first multiprocessing
jsa1 <- .jsap_sa(jsap1, idx = 1) # first SaItem
.jsa_name(jsa1)
mod1 <- .jsa_read(jsa1)
```
